\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-25 paper.qmd
Updated Figure 15

trimesh_removed_pbmc <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c",
               linewidth = 1) +
  geom_point(
    data = umap_pbmc_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.2,
    color = "#000000"
  )  +
  interior_annotation("a1", cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

pbmc_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    #size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("a2", cex = 2)

@@

pbmc_proj_umap_model2 <- ggplot() +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("a3", cex = 2)

@@

trimesh_removed_pbmc_best <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c",
               linewidth = 1) +
  geom_point(
    data = umap_pbmc_scaled_best,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.2,
    color = "#000000"
  )  +
  interior_annotation("b1", cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

pbmc_proj_umap_model1_best <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    #size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("b2", cex = 2)

@@

pbmc_proj_umap_model2_best <- ggplot() +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("b3", cex = 2)

@@

```{r}
#| echo: false
#| fig-cap: ""
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-height: 80%

>>>>

trimesh_removed_pbmc <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  geom_point(
    data = umap_pbmc_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.2,
    color = '#a65628'
  )  +
  interior_annotation("a1", cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

pbmc_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    #size = 0.5,
    alpha = 0.2,
    color = '#a65628') +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("a2", cex = 2)

@@

pbmc_proj_umap_model2 <- ggplot() +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("a3", cex = 2)

@@

trimesh_removed_pbmc_best <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  geom_point(
    data = umap_pbmc_scaled_best,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.3,
    color = "#999999"
  )  +
  interior_annotation("b1", cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

pbmc_proj_umap_model1_best <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    #size = 0.5,
    alpha = 0.2,
    color = "#999999") +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("b2", cex = 2)

@@

pbmc_proj_umap_model2_best <- ggplot() +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("b3", cex = 2)

@@

```{r}
#| echo: false
#| fig-cap: "Model in \\gD{}, on the layout a and c, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$). Layout a shows three-well separated clusters with big separation, while layout c shows close clusters. In $9\\text{-}D$, the data shows close clusters. This is the reason where b1, b2, and b3 shows connected edges between the clusters, and also proved by @fig-pbmc-mse as the best layout, which used the same NLDR method. By viewing the fitted model in $9\\text{-}D$, help to see some unobserved pattern of the data (i) densed points, and (ii) non-linear clusters."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-height: 80%

### 2024-09-25 paper.tex
Updated Figure 15

\caption{\label{fig-model-pbmc-author-proj}}

>>>>

\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the layout
a and c, and two views of the fit in projections from \(9\text{-}D\),
for the PBMC3k data (\(n =  2622\) and \(p = 9\)). Layout a shows
three-well separated clusters with big separation, while layout c shows
close clusters. In \(9\text{-}D\), the data shows close clusters. This
is the reason where b1, b2, and b3 shows connected edges between the
clusters, and also proved by Figure~\ref{fig-pbmc-mse} as the best
layout, which used the same NLDR method. By viewing the fitted model in
\(9\text{-}D\), help to see some unobserved pattern of the data (i)
densed points, and (ii) non-linear clusters.}
